---
title: 'Lab 6: Foundations for statistical inference - Sampling distributions'
author: "Justin Lotz"
date: "26/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#The data

##Load real estate data from the city of Ames, Iowa.
```{r}
download.file("http://www.openintro.org/stat/data/ames.RData", destfile = "ames.RData")
load("ames.RData")
```

##Create short names for above ground living areas and the sales price.
```{r}
area <- ames$Gr.Liv.Area
price <- ames$SalePrice
```

##Histogram of the distribution of area in in our population of home sales and describe the distribution
```{r}
summary(area)
hist(area)

#The distribution is right skewed with the majority of the living spaces being between 1000 and 2000 square feet. 
```

#The unknown sampling distribution

##Create an area sample size of 50 homes in Ames and describe the distribution of the sample. How does the sample distribution compare to the entire population?
```{r}
samp1 <- sample(area, 50)
hist(samp1)

#Similair to the distribution of the entire populaion, the sample area distribution is also right skewed with the majority of the homes having a living area of between 1000 and 2000 square feet. This is to be expected since the sample should represent a snpoashot of the populatioin in its entirity.

```

##Find the average area of homes in Ames by calculating the mean of the sample.
```{r}
mean(samp1)
```

##Create another sample of the Ames home areas and compare the mean to the 1st sample.
```{r}
samp2 <- sample(area, 50)
mean(samp2)

#The sample 2 mean is significantly lower than the first sample.
```

##Create a sample with 100 and 1000 homes and evaluate which sampling mean is most accurate compared to the true population mean of 1499.69 square feet.
```{r}
mean(sample(area, 100))
mean(sample(area, 1000))

#The sample size 0f 1000 will be more accurate. The larger the sample, the closer the approximated mean will be to the full population.
```

##Generate 5000 samples and compute the sample mean of each
```{r}
sample_means50 <- rep(NA, 5000)

for (i in 1:5000) {
samp <- sample(area, 50)
sample_means50[i] <- mean(samp)}


hist(sample_means50, breaks = 25)
```

##How many elements are there in sample_means50? Describe the sampling distribution, and be sure to specifically note its center. Would you expect the distribution to change if we instead collected 50,000 sample means?
```{r}
#There are 5000 elements in sample_means50. The sample is uniformally distributed showing that the sampling distribution of the mean is close to 1500. If we had 50000 sample means I would expect the distribution to be even denser around the actual mean of 1499.
```

#Interlude: The for loop
```{r}
#Initialize a vector of 100 zeros called sample_means_small.
sample_means_small <- rep(NA, 100)

#Run a loop that takes a sample of size 50 from area and stores the sample mean in sample_means_small, but only iterate from 1 to 100.
for (i in 1:100) {
samp11 <- sample(area, 50)
sample_means_small[i] <- mean(samp11)}
```

#Sample size and the sampling distribution
```{r}
sample_means10 <- rep(NA, 5000)
sample_means100 <- rep(NA, 5000)

for(i in 1:5000){
samp <- sample(area, 10)
sample_means10[i] <- mean(samp)
samp <- sample(area, 100)
sample_means100[i] <- mean(samp)
}


```

##Plot the 3 distributions
```{r}
par(mfrow = c(3, 1))

xlimits <- range(sample_means10)

hist(sample_means10, breaks = 20, xlim = xlimits)
hist(sample_means50, breaks = 20, xlim = xlimits)
hist(sample_means100, breaks = 20, xlim = xlimits)
```

##When the sample size is larger, what happens to the center? What about the spread?
```{r}
#When the sample size is larger the center becomes more concentrated and the spread becomes narrower.
```

#On your own

```{r}
#Take a random sample of size 50 from price. Using this sample, what is your best point estimate of the population mean?
sample_price50 <- sample(price, 50)
mean(sample_price50)
##The mean is 174277.6

#Simulate the sampling distribution for ¯xprice by taking 5000 samples from the population of size 50 and computing 5000 sample means.
sample_means50 <- rep(NA, 5000)

for (i in 1:5000) {
samp <- sample(price, 50)
sample_means50[i] <- mean(samp)}

#Plot the data, then describe the shape of this sampling distribution.
hist(sample_means50, breaks = 25)
##The sampling distribution appears to be symmetrical with prices averaging at approximately 180000

#Calculate and report the population mean.
mean(price)
##The price population mean is 180796

#Change your sample size from 50 to 150, then compute the sampling distribution using the same method as above, and store these means in a new vector called sample_means150.

sample_means150 <- rep(NA, 5000)

for (i in 1:5000) {
samp <- sample(price, 150)  
sample_means150[i] <- mean(samp)}

hist(sample_means150, breaks = 25)
##The distribution is similair to the sample of 50, however the center is more concentrated at 180 000.




```




